Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-24 files/Chapter3/femconvergence.tex
chapter 3

From the textbook \cite{OR76}, the following approximation theorem for finite elements is given verbatim:
\newtheorem*{Interpolation}{Theorem}
\begin{Interpolation}[The Interpolation Theorem for Finite Elements] 
	Let $\Omega$ be an open bounded domain in $\mathcal{R}^n$ satisfying the cone condition. Let $k$ be a fixed integer and $m$ an integer such that $0\leq m \leq k+1$. Let $\Pi \in L(H^{k+1}(\Omega), H^{m}(\Omega))$ be such that
	\begin{eqnarray}
	    \Pi u = u \ \ \ \ \textrm{ for all } u \in \mathcal{P}_k(\Omega)
	\end{eqnarray}
	Then for any $u \in H^{k+1}(\Omega)$ and for sufficiently small $h$, there exists positive a constant $C$, independent of $u$ and $h$, such that
	\begin{eqnarray}
	    ||u - \Pi u||_{H^m(\Omega)} \leq C \frac{h^{k+1}}{p^m} |u|_H^{k+1}(\Omega)
	\end{eqnarray}
	where $|u|_H^{k+1}(\Omega)$ is the seminorm.
\end{Interpolation}

@@

The first step to obtain estimates for the eigenvalues, it to show that the elements of $B_m$ are linearly independent. In the first part we follow the approach in \cite{Zie2000} (A Computable Spectral Theorem - Martin Ziegler - 2000). The author introduced the quantity $\mu_m$.

@@

\newtheorem{Cor_1}{Proposition}
\begin{Cor_1}
	$0 \leq \sigma_{m}^{h} < 1$ and $\lambda_{m}^{h} - \lambda_{m} \leq \lambda_{m}^{h}\sigma_{m}^{h}$
\end{Cor_1}
\begin{proof}
	Starting with the result of Lemma \ref{Lem_1},
	\begin{align*}
		\lambda_{m}^{h} &\leq \frac{\lambda_{m}}{\mu_{m}^{h}} \\
		\lambda_{m}^{h}\mu_{m}^{h} &\leq \lambda_{m} \\
		-\lambda_{m} &\leq -\lambda_{m}^{h}\mu_{m}^{h} \\
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h} - \lambda_{m}^{h}\mu_{m}^{h} \\
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h}\sigma_{m}^{h}
	\end{align*}
\end{proof}

@@

\newtheorem{Prop_3}[Prop_1]{Proposition} 
\begin{Prop_3}
	\label{Prop_3}
	$\sigma_{m}^{h} = \max\left\{ 2( u,u-Pu )-||u-Pu||^{2} \ | \ u \in B_{m} \right\}$
\end{Prop_3}
\begin{proof}
	Let $u \in B_{m}$. Then
	\begin{eqnarray*}
		||u - Pu||^{2} &=& c( u - Pu, u - Pu ), \\
						&=& c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ), \\
						&=& 2c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ) - c( u, u ),\\
						&=& 2c( u, u - Pu ) + c( Pu, Pu ) - c( u, u ).\\
	\end{eqnarray*}
	consequently
	\begin{eqnarray*}
		c( u, u ) - c( Pu, Pu )  & = & 2c( u, u - Pu ) - ||u - Pu||^{2}.
	\end{eqnarray*}

@@

\newtheorem{Lem_2}[Lem_1]{Lemma} \label{Lem_2}
\begin{Lem_2}
	$\sigma_{m}^{h} \leq \max \left\{2||u^{*}- Pu^{*}|| ||u-Pu|| \ | \ u \in B_{m} \right\}$
\end{Lem_2}
\begin{proof}
	From Proposition \ref{Prop_3},
	\begin{eqnarray*}
		\sigma_{m}^{h} & = & \max\left\{ 2c( u,u-Pu )-||u-Pu||^{2} \ | \ u \in B_{m} \right\}, \\
					& \leq & \max\left\{ 2c( u,u-Pu ) \ | \ u \in B_{m} \right\}. \\
	\end{eqnarray*}

@@

\newtheorem{Prop_5}[Prop_1]{Proposition} \label{Prop_5}
\begin{Prop_5}
	For any $\epsilon > 0$, there exists a $\delta >0$ such that for $h<\delta$,
	\begin{eqnarray*}
	||u^{*} - Pu^{*}|| < \epsilon \ \textrm{ for each } u \in B_{m}\\
	||u - Pu|| < \epsilon \ \textrm{ for each } u \in B_{m}
	\end{eqnarray*}
\end{Prop_5}
\begin{proof}
	Consider a set of basis functions $\Phi \subset S^h$ such that $\phi^(k) = 0$ for each $\phi \in \Phi$.

@@

\newtheorem{Lem_5}[Lem_1]{Lemma} \label{Lem_5}
\begin{Lem_5}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h} ) + \lambda_{m}^{h} - \lambda_{m}
		\end{eqnarray*}
\end{Lem_5}
\begin{proof}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& b(u_{m},u_{m}) - 2b(u_{m},u^{h}_{m}) + b(u^{h}_{m},u^{h}_{m}) \\
										&=& \lambda_{m} c( u_{m}, u_{m} ) - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h}c( u_{m}^{h},u_{m}^{h} )\\
										&=&  \lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h} \\
										&=& 2\lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h} - \lambda_{m}\\
										&=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) + \lambda_{m}^{h} - \lambda_{m}
		\end{eqnarray*}
\end{proof}

@@

\newtheorem{Lem_6}[Lem_1]{Lemma} \label{Lem_6}
\begin{Lem_6}
	For all m and j
	\begin{eqnarray*}
	(\lambda_{j}^{h} - \lambda_{m}) c( Pu_{m}, u_{j}^{h}) = \lambda_{m} c( u_{m}-Pu_{m},u_{j}^{h} )
	\end{eqnarray*}
\end{Lem_6}
\begin{proof}
	Since the term $\lambda_{m}c( Pu,u^{h}_{j})$ appears on both sides of the equation, it is only required to show that
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& \lambda_{m} c( u, u_{j}^{h} ).	
	\end{eqnarray*}
	Since both $u$ and $u_{j}^{h}$ are eigenfunctions, then
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& b(Pu,u_{j}^{h})\\
		\lambda_{m} c( u, u_{j}^{h} ) &=& b(u,u_{j}^{h})
	\end{eqnarray*}

@@

\newtheorem{Lem_7}[Lem_1]{Lemma} \label{Lem_7}
\begin{Lem_7}
	\begin{eqnarray*}
		||Pu - \beta Pu^h_m||^{2} &\leq & {\rho}^{2} ||u_{m} - Pu_{m}||^{2}
	\end{eqnarray*}
\end{Lem_7}
\begin{comment}
\begin{proof}

@@

\newtheorem{Lem_8}[Lem_1]{Lemma} \label{Lem_8}
\begin{Lem_8}
	\begin{eqnarray*}
		||u_{m} - \beta u_{m}^{h}|| &\leq & \left(1+\rho\right)||u_{m}-Pu_{m}||
	\end{eqnarray*}
\end{Lem_8}
\begin{comment}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - \beta u_{m}^{h}|| & \leq & ||u_{m}-Pu_{m}|| + ||Pu_{m} - \beta u_{m}^{h}|| \\
				& \leq & \left(1+\rho\right)||u_{m}-Pu_{m}|| \ \ \textrm{ (by Lemma \ref{Lem_7})}
	\end{eqnarray*}
\end{proof}
\end{comment}

@@

\newtheorem{Lem_9}[Lem_1]{Lemma} \label{Lem_9}
\begin{Lem_9}
	\begin{eqnarray*}
		||u_{m} -  u_{m}^{h}|| &\leq & 2||u_{m}-\beta u^{h}_{m}||
	\end{eqnarray*}
\end{Lem_9}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - u_{m}^{h}|| &=& ||u_{m} - \beta u_{m}^{h} + \beta u_{m}^{h} - u_{m}^{h}||, \\
						& \leq & ||u_{m} - \beta u_{m}^{h}|| + ||\beta u_{m}^{h} - u_{m}^{h}||, \\
						& = & 2||u_{m} - \beta u_{m}^{h}||.	
	\end{eqnarray*}
\end{proof}

>>>>

From the textbook \cite{OR76}, following approximation theorem for finite elements is given verbatim:
\newtheorem*{Interpolation}{Theorem}
\begin{Interpolation}[The Interpolation Theorem for Finite Elements] 
	Let $\Omega$ be an open bounded domain in $\mathcal{R}^n$ satisfying the cone condition. Let $k$ be a fixed integer and $m$ an integer such that $0\leq m \leq k+1$. Let $\Pi \in L(H^{k+1}(\Omega), H^{m}(\Omega))$ be such that
	\begin{eqnarray}
	    \Pi u = u \ \ \ \ \textrm{ for all } u \in \mathcal{P}_k(\Omega)
	\end{eqnarray}
	Then for any $u \in H^{k+1}(\Omega)$ and for sufficiently small $h$, there exists positive a constant $C$, independent of $u$ and $h$, such that
	\begin{eqnarray}
	    ||u - \Pi u||_{H^m(\Omega)} \leq C \frac{h^{k+1}}{p^m} |u|_H^{k+1}(\Omega)
	\end{eqnarray}
	where $|u|_H^{k+1}(\Omega)$ is the seminorm.
\end{Interpolation}

@@

The first step to obtain estimates for the eigenvalues, it to show that the elements of $B_m$ are linearly independent. In the first part we follow the approach in \cite{Zie2000} (A Computable Spectral Theorem - Martin Ziegler - 2000 ?). The author introduced the quantity $\mu_m$.

@@

\newtheorem{Cor_1}[Prop_1]{Proposition}
\begin{Cor_1}
	$0 \leq \sigma_{m}^{h} < 1$ and $\lambda_{m}^{h} - \lambda_{m} \leq \lambda_{m}^{h}\sigma_{m}^{h}$
\end{Cor_1}
\begin{proof}
	Starting with the result of Lemma \ref{Lem_1},
	\begin{align*}
		\lambda_{m}^{h} &\leq \frac{\lambda_{m}}{\mu_{m}^{h}}, \\
		\lambda_{m}^{h}\mu_{m}^{h} &\leq \lambda_{m}. \\
	\end{align*}
	This is equivalent to
	\begin{align*}
		-\lambda_{m} &\leq -\lambda_{m}^{h}\mu_{m}^{h}. 
	\end{align*}
	Add $\lambda_{m}^{h}$ to both sides
	\begin{align*}
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h} - \lambda_{m}^{h}\mu_{m}^{h}. 
	\end{align*}
	And finally using the error bound $\sigma_{m}^{h} = 1 - \mu_{m}^{h}$,
	\begin{align*}
		\lambda_{m}^{h} - \lambda_{m} &\leq \lambda_{m}^{h}\sigma_{m}^{h}.
	\end{align*}
\end{proof}

@@

\newtheorem{Prop_3}[Prop_1]{Proposition} 
\begin{Prop_3}
	\label{Prop_3}
	$\sigma_{m}^{h} = \max\left\{ 2c( u,u-Pu )-||u-Pu||^{2} \ | \ u \in B_{m} \right\}$
\end{Prop_3}
\begin{proof}
	Let $u \in B_{m}$. Then
	\begin{eqnarray*}
		||u - Pu||^{2} &=& c( u - Pu, u - Pu ), \\
						&=& c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ), \\
						&=& 2c( u, u ) - 2 c( u, Pu ) + c( Pu, Pu ) - c( u, u ),\\
						&=& 2c( u, u - Pu ) + c( Pu, Pu ) - c( u, u ).\\
	\end{eqnarray*}
	consequently
	\begin{eqnarray*}
		c( u, u ) - c( Pu, Pu )  & = & 2c( u, u - Pu ) - ||u - Pu||^{2}.
	\end{eqnarray*}

@@

\newtheorem{Lem_2}[Lem_1]{Lemma} 
\begin{Lem_2}
	\label{Lem_2}
	$\sigma_{m}^{h} \leq \max \left\{2||u^{*}- Pu^{*}|| ||u-Pu|| \ | \ u \in B_{m} \right\}$
\end{Lem_2}
\begin{proof}
	From Proposition \ref{Prop_3},
	\begin{eqnarray*}
		\sigma_{m}^{h} & = & \max\left\{ 2c( u,u-Pu )-||u-Pu||^{2} \ | \ u \in B_{m} \right\}, \\
					& \leq & \max\left\{ 2c( u,u-Pu ) \ | \ u \in B_{m} \right\}. \\
	\end{eqnarray*}

@@

\newtheorem{Prop_5}[Prop_1]{Proposition} 
\begin{Prop_5}
	\label{Prop_5}
	For any $\epsilon > 0$, there exists a $\delta >0$ such that for $h<\delta$,
	\begin{eqnarray*}
	||u^{*} - Pu^{*}|| < \epsilon \ \textrm{ for each } u \in B_{m}\\
	||u - Pu|| < \epsilon \ \textrm{ for each } u \in B_{m}
	\end{eqnarray*}
\end{Prop_5}
\begin{proof}
	Consider a set of basis functions $\Phi \subset S^h$ such that $\phi^(k) = 0$ for each $\phi \in \Phi$.

@@

\newtheorem{Lem_5}[Lem_1]{Lemma} 
\begin{Lem_5}
	\label{Lem_5}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h} ) + \lambda_{m}^{h} - \lambda_{m}
		\end{eqnarray*}
\end{Lem_5}
\begin{proof}
	\begin{eqnarray*}
		b(u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) &=& b(u_{m},u_{m}) - 2b(u_{m},u^{h}_{m}) + b(u^{h}_{m},u^{h}_{m}) \\
										&=& \lambda_{m} c( u_{m}, u_{m} ) - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h}c( u_{m}^{h},u_{m}^{h} )\\
										&=&  \lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h} \\
										&=& 2\lambda_{m} - 2\lambda_{m} c( u_{m}, u^{h}_{m} ) + \lambda_{m}^{h} - \lambda_{m}\\
										&=& \lambda_{m}c( u_{m}-u_{m}^{h},u_{m}-u_{m}^{h}) + \lambda_{m}^{h} - \lambda_{m}
		\end{eqnarray*}
\end{proof}

@@

\newtheorem{Lem_6}[Lem_1]{Lemma} 
\begin{Lem_6}
	\label{Lem_6}
	For all m and j
	\begin{eqnarray*}
	(\lambda_{j}^{h} - \lambda_{m}) c( Pu_{m}, u_{j}^{h}) = \lambda_{m} c( u_{m}-Pu_{m},u_{j}^{h} )
	\end{eqnarray*}
\end{Lem_6}
\begin{proof}
	Since the term $\lambda_{m}c( Pu,u^{h}_{j})$ appears on both sides of the equation, it is only required to show that
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& \lambda_{m} c( u, u_{j}^{h} ).	
	\end{eqnarray*}
	Since both $u$ and $u_{j}^{h}$ are eigenfunctions, then
	\begin{eqnarray*}
		\lambda_{j}^{h}c( Pu, u_{j}^{h} ) &=& b(Pu,u_{j}^{h})\\
		\lambda_{m} c( u, u_{j}^{h} ) &=& b(u,u_{j}^{h})
	\end{eqnarray*}

@@

\newtheorem{Lem_7}[Lem_1]{Lemma} 
\begin{Lem_7}
	\label{Lem_7}
	\begin{eqnarray*}
		||Pu - \beta Pu^h_m||^{2} &\leq & {\rho}^{2} ||u_{m} - Pu_{m}||^{2}
	\end{eqnarray*}
\end{Lem_7}
\begin{comment}
\begin{proof}

@@

\newtheorem{Lem_8}[Lem_1]{Lemma} 
\begin{Lem_8}
	\label{Lem_8}
	\begin{eqnarray*}
		||u_{m} - \beta u_{m}^{h}|| &\leq & \left(1+\rho\right)||u_{m}-Pu_{m}||
	\end{eqnarray*}
\end{Lem_8}
\begin{comment}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - \beta u_{m}^{h}|| & \leq & ||u_{m}-Pu_{m}|| + ||Pu_{m} - \beta u_{m}^{h}|| \\
				& \leq & \left(1+\rho\right)||u_{m}-Pu_{m}|| \ \ \textrm{ (by Lemma \ref{Lem_7})}
	\end{eqnarray*}
\end{proof}
\end{comment}

@@

\newtheorem{Lem_9}[Lem_1]{Lemma} 
\begin{Lem_9}
	\label{Lem_9}
	\begin{eqnarray*}
		||u_{m} -  u_{m}^{h}|| &\leq & 2||u_{m}-\beta u^{h}_{m}||
	\end{eqnarray*}
\end{Lem_9}
\begin{proof}
	\begin{eqnarray*}
	||u_{m} - u_{m}^{h}|| &=& ||u_{m} - \beta u_{m}^{h} + \beta u_{m}^{h} - u_{m}^{h}||, \\
						& \leq & ||u_{m} - \beta u_{m}^{h}|| + ||\beta u_{m}^{h} - u_{m}^{h}||, \\
						& = & 2||u_{m} - \beta u_{m}^{h}||.	
	\end{eqnarray*}
\end{proof}
